\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-05-23 paper/paper.tex
Updates to Priors, updated pi_compare

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

@@

Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution. We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that wearout implies an increasing hazard function. The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data. For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.

@@

\begin{enumerate}
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\end{enumerate}

@@

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive brand models. We found that there was insufficient information in the data to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of the various ${\pi_{d}$, the probability of infant mortality for drive model $d$.

\subsection{Priors}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior. For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors. As for the location hyper-parameters, we select weakly informative priors consistent with our prior information on hard-drives. The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification? Should maybe be lower.) For $\eta_{\sigma, 2}$, 95\% of prior mass is on values greater than $0.037$. Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$. The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.

@@

\subsection{Computation}
Each model was fit in {\t R} \nocite{r} using Hamiltonian Monte Carlo
\cite{betancourt} as implemented in {\tt
  rstan}\cite{rstan}. Multiple chain were run with warmup iterations. Convergence was assessed by examining posterior
plots and checking that Gelman-Rubin diagnostics were close to 1.

\section{Data analysis}
\subsection{Results}
The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit. The upper left panel shows the estimated failure curve fit to all data ignoring drive model as a factor (Model 1). The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.

@@

Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.

\subsection{Model Comparisons}
First look at the overlaid posterior credible bands.  Perhaps pick 4 different drive models and present a plot like the figure below.  Bands are for full model, and the solid line is the full model fit. \\

@@

\begin{figure}[H]
    \centering
   \includegraphics[width=5.0in]{fig/pi_bymod.pdf}
		\caption{$\pi$ for All Drive Models \label{fig:first}} 
\end{figure}

\section{Concluding Remarks and Extensions}
Review the advantages of fitting the GFLP model and offer future ideas.

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage[numbers]{natbib}
\usepackage{url} % not crucial - just used below for the URL

@@

Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution. We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that,by nature, wearout produces an increasing hazard function. The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often, as is true of the data presented here, features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data. For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.

@@

\begin{enumerate}
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\end{enumerate}

@@

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive brand models. We found that there was insufficient information in the data to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).

\subsection{Priors}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.

For models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:

\begin{enumerate}
\item This model constrains all drive models to the same failure distribution. For this ``reduced" model we assume the priors :

$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad t_{0.5,1} \sim \op{Normal}(8,4)$$
$$\sigma_2 \sim \op{Lognormal}(0, 2.5) \quad t_{0.2,2} \sim \op{N}(10,4)$$

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$. The prior on $\sigma_1$ and $\sigma_2$ put it on the interval $(0.02, 61.1)$ with $90\%$ probability. (Note that, due to symmetry, the Weibull shape parameter has the same prior interval and that a value of 1, corresponding to a constant hazard, is the median of this prior distribuion.) $t_{0.5,1}$ has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ it is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.

\item We allow $t_{pd}$ to vary by drive-model. To help with model identifiability, we tighten the priors on the defective mode:
$$ \mbox{logit}^{-1}\pi \sim \op{Normal}(-3,1),\quad \sigma_1 \sim \op{Lognormal}(0, 1), \quad t_{p1} \sim \op{Normal}(7,2),$$
implying now a prior $90\%$ probability that $\sigma_1$ is between $0.19$ and $5.18$ and $t_{0.5,1}$ between $3.7$ and $10.3$, which, translating from the log-scale, is equivalent to an interval from 1.5 days to 3.4 years.

\item $\sigma_{d2}$ is allowed to vary. Priors for constrained parameters are the same as for model 2.

\item $\pi_d$ varies. Priors for constrained parameters are the same as for model 2.

\end{enumerate}

For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors. As for the location hyper-parameters, we select weakly informative priors consistent with our prior information on hard-drives. The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification? Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$. Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$. The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.

@@

\subsection{Computation}
Each model was fit using the {\tt
  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\cite{betancourt}. Multiple chain were run for 1500 iterations after 1500 warmup iterations. Convergence was assessed by examining posterior
plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for models 1,2 and 3 and 16 chains were run for model 4.

\section{Data analysis}

\subsection{Model Comparisons}
The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit. The upper left panel shows the estimated failure curve fit to all data ignoring drive model as a factor (Model 1). The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.

@@

Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward. 
First look at the overlaid posterior credible bands.  Perhaps pick 4 different drive models and present a plot like the figure below.  Bands are for full model, and the solid line is the full model fit. \\

@@

\begin{figure}[H]
    \centering
   \includegraphics[width=5.0in]{fig/pi_compare.pdf}
		\caption{$\pi$ for All Drive Models \label{fig:first}} 
\end{figure}

\section{Discussion}
Review the advantages of fitting the GFLP model and offer future ideas.

Advances in computer technology have made computationally intensive Bayesian methods for fitting hierarchical models feasible in practice. Analysis using HMC is accessible to practitioners by Stan. HMC can be very efficient in terms of effective samples because of the way that it uses joint updates for all parameters using Hamiltonian dynamics to rapidly explore the target distribution. Hierarchical modeling can be a useful way to pool information to estimate large models. By using Bayesian methods like the one presented here we can avoid using potentially problematic asymptotic approximations, since all posterior uncertainty is automatically integrated into averages computed from functions of the samples.
